### 2016-09-20 Thesis/DCG_sb-wi-thesis-template/MSpC/MSpC.tex
proteasome additions final edits after erin

For example, reasonably accurate concentration plots were obtained for the PAF ($\alpha$6) and PBD ($\beta$4) subunits that are encoded by the PAF1/2 and PBF1/2 gene pairs, and for the PAG ($\alpha$7) and PBF ($\beta$6) subunits that are encoded by single PAG1 and PBF1 genes (R$\textsuperscript{2}$  from 0.94 to 0.99).
Even when we calculated uNSAF values for individual isoforms added to the \textit{E. coli} lysate, strong linear responses were obtained (e.g., the PAF1/PAF2 and PBD1/PBD2 pairs) with robust correlations (R$\textsuperscript{2}$ from 0.89 to 0.95) (Figure {fig:proteasomespike} D).
Taken together, MSpC worked well for relative LFQ analysis of a multi-subunit complex and its individual subunits and isoforms within a complex proteomic mixture.

@@

Samples were analyzed by ultra-high performance liquid chromatography (UPLC) (nanoAcquity, Waters Corporation) connected online to an electrospray ionization LTQ Velos Orbitrap mass spectrometer (Thermo Scientific).  Separation employed a 100 x 365 $\mu$m fused silica capillary micro-column packed with 20 cm of 1.7 $\mu$m-diameter, 130-$\AA$ pore size, C18 beads (Waters BEH), with an emitter tip pulled to approximately 1 $\mu$m using a laser puller (Sutter Instruments).  Peptides were loaded at a flow-rate of 400 nL/min for 30 min and then eluted over 120 min at a flow-rate of 300 nL/min with a 2\% to 30\% acetonitrile gradient in 0.1\% formic acid.  Full-mass scans were performed in the FT Orbitrap with a mass range of 300-1500 m/z at a resolution of 60,000, followed by ten MS/MS high energy C-trap dissociation scans of the ten highest intensity parent ions at 42\% normalized collision energy and 7,500 resolution, with a mass range starting at 100 m/z.  Dynamic exclusion was enabled with a repeat count of two over the duration of 30 sec and an exclusion window of 120 sec.

@@

Protein identifications were determined using the Morpheus search engine \citep{wenger13}.  Raw data was searched with the Thermo module of Morpheus revision 151 downloaded and compiled from source code available at \url{http://sourceforge.net/projects/morpheus-ms/} using Microsoft Visual Studio 2013 professional edition.  The following parameters were used to search all databases: unknown precursor charge states - +2, +3, +4; maximum number of MS/MS peaks = 400; assign charge states - enabled; de-isotope - disabled; generate target decoy database on the fly; protease trypsin (no proline rule); maximum missed cleavages = 2; initiator methionine behavior - variable; fixed modification of carbamidomethylation of cysteine; variable modification of oxidation of methionine; maximum variable modification isoforms per peptide = 1024; precursor mass tolerance = $\pm$ 2.1 Da monoisotopic (recommended parameters to account for neutral loss); precursor monoisotopic peak correction - disabled; product mass tolerance = $\pm$ 0.025 Da monoisotopic; consider modified forms as unique peptides - false; maximum threads = 8; minimize memory usage - false.   
MSpC quantification of Universal Proteome Standard 2 (UPS2) protein sequences exploited the MS/MS analysis of individual USP2 proteins (Sigma-Aldrich) mixed at various concentrations with egg or embryo extracts from \textit{Xenopus laevis} available in the PRoteomics IDEntifications (PRIDE) repository \citep{vizcaino13} using identifier - PXD000902 and the available proteomics database - pita\_v1.71.protein.name.fa.  For our analysis, both the raw MS/MS data and resulting FASTA files for the USP2 and egg and embryo proteomes were obtained from PRIDE.  The database used for searching the MS/MS data of \textit{Arabidopsis} proteasomes spiked into total \textit{E. coli} peptides was generated by combining Uniprot K12 \textit{E. coli} reference proteome UP000000625 with a common contaminant database, and then mixing the merged dataset with FASTA sequences for all proteoforms of all known proteasome subunits and associated proteins \citep{book10} obtained from the TAIR10\_pep\_20101214 FASTA database available within The \textit{Arabidopsis} Information Resource (TAIR) version 10.  All FASTA files are available for download in the Supporting Information.  The datasets were analyzed by MSpC with a 1\% PSM false discovery rate (FDR) and a 1\% protein group FDR to determine NSAF, dNSAF, and uNSAF values for each protein group. Two separate analyses were performed in which one unique, or alternatively two unique peptides were required to quantify a protein group.  NSAF values were calculated according to formulas 1a, 2a, and 3a from Figure 1 of Zhang et al. \citep{zhang10}.

@@

\section {Discussion on Requiring Two Unique Peptides to Quantify a Protein}
\label{sec:discussiontwopep}
Occasionally, some researchers may want to use a more stringent criterion for quantification such as requiring a protein to have more than one unique identifying peptide. To see how this might affect our data analysis, we re-analyzed our results shown in Figure \ref{fig:UPS2twopep}, this time requiring two unique peptides to quantify an individual protein. The results point to a very small increase in linearity observed in the average plots; however, there is a slight decrease in linearity for the egg sample (0.886 to 0.865) and a larger decrease in linearity for the individual UPS2 protein plot for the embryo sample (0.827 to 0.723). The decrease in linearity in the embryo sample is due to the analysis removing a low abundance UPS2 protein (O00762ups) identified with only one unique peptide. While some have suggested that requiring more than one unique peptide to identify a protein is an ideal approach, requiring two peptides for identifications in database searches reduces the number of protein identifications in the target database more than those in the decoy database and results in increased false discovery rates . While we recognize that researchers may want to implement more stringent requirements than what is typically used in database searching to quantify a set of proteins, there are two cases where requiring two unique peptides may not be ideal in a quantitative analysis. Firstly, low abundance proteins that have few PSMs might be identified by only a single peptide and thus be erroneously thrown out of the analysis. Secondly, there may be only one unique peptide that can differentiate between families of homologous proteins. In this second case, requiring two unique peptides would remove these homologous proteins from the MSpC analysis, even if they had a large number of PSMs. Because of these reasons and because of the decreased linearity observed when requiring proteins to have two unique peptides (Figure \ref{fig:UPS2twopep}) as compared to one unique peptide (Figure \ref{fig:UPS2}), we suggest caution in requiring more than one unique peptide per protein.

>>>>

For example, reasonably accurate concentration plots were obtained for the PAF ($\alpha$6) and PBD ($\beta$4) subunits that are encoded by the PAF1/2 and PBF1/2 gene pairs, and for the PAG ($\alpha$7) and PBF ($\beta$6) subunits that are encoded by single PAG1 and PBF1 genes (R$\textsuperscript{2}$  from 0.94 to 0.99).
Even when we calculated uNSAF values for individual isoforms added to the \textit{E. coli} lysate, strong linear responses were obtained (e.g., the PAF1/PAF2 and PBD1/PBD2 pairs) with robust correlations (R$\textsuperscript{2}$ from 0.89 to 0.95) (Figure \ref{fig:proteasomespike} D).
Taken together, MSpC worked well for relative LFQ analysis of a multi-subunit complex and its individual subunits and isoforms within a complex proteomic mixture.

@@

Samples were analyzed by ultra-high performance liquid chromatography (UPLC) (nanoAcquity, Waters Corporation) connected online to an electrospray ionization LTQ Velos Orbitrap mass spectrometer (Thermo Scientific).  Separation employed a 100 x 365 $\mu$m fused silica capillary micro-column packed with 20 cm of 1.7 $\mu$m-diameter, 130-Å pore size, C18 beads (Waters BEH), with an emitter tip pulled to approximately 1 $\mu$m using a laser puller (Sutter Instruments).  Peptides were loaded at a flow-rate of 400 nL/min for 30 min and then eluted over 120 min at a flow-rate of 300 nL/min with a 2\% to 30\% acetonitrile gradient in 0.1\% formic acid.  Full-mass scans were performed in the FT Orbitrap with a mass range of 300-1500 \textit{m/z} at a resolution of 60,000, followed by ten MS/MS high energy C-trap dissociation scans of the ten highest intensity parent ions at 42\% normalized collision energy and 7,500 resolution, with a mass range starting at 100 \textit{m/z}.  Dynamic exclusion was enabled with a repeat count of two over the duration of 30 s and an exclusion window of 120 s.

@@

Protein identifications were determined using the Morpheus search engine \citep{wenger13}.  Raw data was searched with the Thermo module of Morpheus revision 151 downloaded and compiled from source code available at \url{http://sourceforge.net/projects/morpheus-ms/} using Microsoft Visual Studio 2013 professional edition.  The following parameters were used to search all databases: unknown precursor charge states - +2, +3, +4; maximum number of MS/MS peaks = 400; assign charge states - enabled; de-isotope - disabled; generate target decoy database on the fly; protease trypsin (no proline rule); maximum missed cleavages = 2; initiator methionine behavior - variable; fixed modification of carbamidomethylation of cysteine; variable modification of oxidation of methionine; maximum variable modification isoforms per peptide = 1024; precursor mass tolerance = $\pm$ 2.1 Da monoisotopic (recommended parameters to account for neutral loss); precursor monoisotopic peak correction - disabled; product mass tolerance = $\pm$ 0.025 Da monoisotopic; consider modified forms as unique peptides - false; maximum threads = 8; minimize memory usage - false.   
MSpC quantification of Universal Proteome Standard 2 (UPS2) protein sequences exploited the MS/MS analysis of individual USP2 proteins (Sigma-Aldrich) mixed at various concentrations with egg or embryo extracts from \textit{Xenopus laevis} available in the PRoteomics IDEntifications (PRIDE) repository \citep{vizcaino13} using identifier - PXD000902 and the available proteomics database - pita\_v1.71.protein.name.fa.  For our analysis, both the raw MS/MS data and resulting FASTA files for the USP2 and egg and embryo proteomes were obtained from PRIDE.  The database used for searching the MS/MS data of \textit{Arabidopsis} proteasomes spiked into total \textit{E. coli} peptides was generated by combining Uniprot K12 \textit{E. coli} reference proteome UP000000625 with a common contaminant database, and then mixing the merged dataset with FASTA sequences for all proteoforms of all known proteasome subunits and associated proteins \citep{book10} obtained from the TAIR10\_pep\_20101214 FASTA database available within The \textit{Arabidopsis} Information Resource (TAIR) version 10.  All FASTA files are available for download in the Supporting Information.  The datasets were analyzed by MSpC with a 1\% PSM false discovery rate (FDR) and a 1\% protein group FDR to determine NSAF, dNSAF, and uNSAF values for each protein group. Two separate analyses were performed in which one unique, or alternatively two unique peptides were required to quantify a protein group.  NSAF values were calculated according to formulas 1a, 2a, and 3a from Figure 1 of Zhang \textit{et al}. \citep{zhang10}.

@@

\section {Discussion on Requiring Two Unique Peptides to Quantify a Protein}
\label{sec:discussiontwopep}
Occasionally, some researchers may want to use a more stringent criterion for quantification such as requiring a protein to have more than one unique identifying peptide. To see how this might affect our data analysis, we re-analyzed our results shown in Figure \ref{fig:UPS2twopep}, this time requiring two unique peptides to quantify an individual protein. The results point to a very small increase in linearity observed in the average plots; however, there is a slight decrease in linearity for the egg sample (0.886 to 0.865) and a larger decrease in linearity for the individual UPS2 protein plot for the embryo sample (0.827 to 0.723). The decrease in linearity in the embryo sample is due to the analysis removing a low abundance UPS2 protein (O00762ups) identified with only one unique peptide. While some have suggested that requiring more than one unique peptide to identify a protein is an ideal approach, requiring two peptides for identifications in database searches reduces the number of protein identifications in the target database more than those in the decoy database and results in increased false discovery rates \citep{gupta09}. While we recognize that researchers may want to implement more stringent requirements than what is typically used in database searching to quantify a set of proteins, there are two cases where requiring two unique peptides may not be ideal in a quantitative analysis. Firstly, low abundance proteins that have few PSMs might be identified by only a single peptide and thus be erroneously thrown out of the analysis. Secondly, there may be only one unique peptide that can differentiate between families of homologous proteins. In this second case, requiring two unique peptides would remove these homologous proteins from the MSpC analysis, even if they had a large number of PSMs. Because of these reasons and because of the decreased linearity observed when requiring proteins to have two unique peptides (Figure \ref{fig:UPS2twopep}) as compared to one unique peptide (Figure \ref{fig:UPS2}), we suggest caution in requiring more than one unique peptide per protein.
